BumpyMatrix: a Bioconductor package to represent a 2-dimensional array of non-scalar objects

### 2023-01-19 paper.Rmd
Added some basic references.

Many bioinformatics workflows in the R/Bioconductor ecosystem are based on a 2-dimensional data representation where one dimension corresponds to features and the other dimension corresponds to samples.
Obvious examples of this paradigm include gene-by-sample expression matrices for transcriptome profiling experiments, peak-by-sample matrices from ChIP-seq experiments, and gene-by-cell count matrices for single-cell RNA-seq experiments.
A more sophisticated example is the `SummarizedExperiment` class, a data structure that contains any number of assay matrices, feature- and sample-level annotations as well as any additional metadata for a dataset.
The `SummarizedExperiment` is a popular representation for genomics data, serving as the cornerstone of a wide variety of analysis pipelines ranging from RNA sequencing, epigenomics, single-cell analyses and more.

However, not all types of experimental data can be easily inserted into a `SummarizedExperiment`.
The use of assay matrices assumes that each feature/sample combination has one experimental observation (or zero observations, if missing values are allowed in the matrix).
It is more difficult to store data types where each feature/sample combination is associated with multiple observations.
Our motivating example is that of a highly multiplexed RNA fluorescence _in situ_ hybridization (RNA-FISH) experiment.
Here, the features are genes, the samples are cells, and each gene/cell combination is associated with spatial coordinates for zero, one or multiple transcript molecules.

The RNA-FISH transcript coordinate data has no obvious location inside the `SummarizedExperiment`.
We cannot store it in the assays as it cannot be represented as an ordinary R matrix. 
We might consider grouping the coordinates by cell and storing them in the sample-level annotations, but this would mean that any feature-level slicing of the `SummarizedExperiment` would not be propagated to the transcript coordinates.
(And vice versa for storing them in the feature-level annotations, whereupon they would not respond to any sample-level slicing.)
This negates one of the major advantages of the `SummarizedExperiment` concept - that any operations applied to a `SummarizedExperiment` will propagate across all assays and annotations contained within,
ensuring synchronization and avoiding common book-keeping errors throughout the course of an analysis.

>>>>

Many bioinformatics workflows in the R/Bioconductor ecosystem are based on a 2-dimensional data representation where one dimension corresponds to features and the other dimension corresponds to samples.
Obvious examples of this paradigm include gene-by-sample expression matrices for transcriptome profiling experiments, peak-by-sample matrices from ChIP-seq experiments, and gene-by-cell count matrices for single-cell RNA-seq experiments.
A more sophisticated example is the `SummarizedExperiment` class [@huber2015orchestrating], a data structure that contains any number of assay matrices, feature- and sample-level annotations as well as any additional metadata for a dataset.
The `SummarizedExperiment` is a popular representation for genomics data, serving as the cornerstone of a wide variety of analysis pipelines ranging from RNA sequencing, epigenomics, single-cell analyses and more.

However, not all types of experimental data can be easily inserted into a `SummarizedExperiment`.
The use of assay matrices assumes that each feature/sample combination has one experimental observation (or zero observations, if missing values are allowed in the matrix).
It is more difficult to store data types where each feature/sample combination is associated with multiple observations.
Our motivating example is that of a highly multiplexed RNA fluorescence _in situ_ hybridization (RNA-FISH) experiment [@moffitt2016highthroughput;@eng2019transcriptome]. 
Here, the features are genes, the samples are cells, and each gene/cell combination is associated with spatial coordinates for zero, one or multiple transcript molecules.

The RNA-FISH transcript coordinate data has no obvious location inside the `SummarizedExperiment`.
We cannot store it in the assays as it cannot be represented as an ordinary R matrix. 
We might consider grouping the coordinates by cell and storing them in the sample-level annotations, but this would mean that any feature-level slicing of the `SummarizedExperiment` would not be propagated to the transcript coordinates.
(And vice versa for storing them in the feature-level annotations, whereupon they would not respond to any sample-level slicing.)
This negates one of the major advantages of the `SummarizedExperiment` concept - that any operations applied to a `SummarizedExperiment` will propagate across all assays and annotations contained within,
ensuring synchronization and avoiding common book-keeping errors throughout the course of an analysis, e.g., using the single-cell workflow described by @amezquita2020orchestrating.
